Accessing the nucleon transverse structure in inclusive deep inelastic scattering

### 2017-06-06 1607 - JMC for g2/Paper/JMC_dis.tex
added preprint no.'
g2+g1 definition attempt - see e-mail from June 6

\preprint{JLAB-THY-17-????}

@@

In this paper, we discuss a novel way of accessing the transversity parton distribution function (PDF) and measuring the proton's tensor charge in totally inclusive Deep Inelastic Scattering. 
We revisit the standard analysis of the DIS handbag diagram, taking into account the fact that on-shell quarks cannot, in fact, be present in the final state, but they rather decay and form jets of hadrons. This is sufficient to modify the structure of the DIS
cut diagram, even if none of the hadrons is detected in the final
state. For a proper description of this effect, we include ``jet
correlators'' into the analysis, and pay particular attention to ensuring that our results are gauge invariant.

@@

The correlator $\Xi$ can be parametrized in terms of jet parton correlation
functions $A_i$ and $B_i$ through a Lorentz covariant Dirac decomposition that utilizes the vectors $l$ and $n_+$, 
\begin{equation}
\Xi(l,n_+) = \Lambda A_1(l^2)\,\id + A_2(l^2)\,\lslash 
+ \frac{\Lambda^2}{l \cdot n_{+}} \nslash_+ \, B_{1}(l^2)
+ \frac{i \Lambda}{2 l \cdot n_{+}} [\,\lslash,\nslash_+ ] \, B_{2}(l^2) \ ,
\label{e:jetexpansion}
\end{equation} 
where $\Lambda$ is an arbitrary scale, introduced for power counting purposes.
In this parametrization, no terms proportional to $\gamma_5$ enter because of parity invariance. Time reversal invariance in QCD requires $B_{2}=0$, while $B_{1}$ contributes only at twist-4 order and will not be considered further in this paper. We focus, instead, on the role of chiral odd terms in the $g_2$ structure function up to twist 3. At this order, 
\begin{equation}
  \Xi(l,n_+) = \Lambda A_1(l^2)\,\id + A_2(l^2)\,\lslash 
    + {\cal O}(\Lambda^2/Q^2)
\label{e:jetexpansion-tw3}
\end{equation} 
is nothing else than the cut quark propagator; note however, that we consider
here the full QCD vacuum rather than the perturbative one (or, in other words, the interacting rather than the free quark fields).
The $A_1$ and $A_2$ terms can be interpreted in terms of the spectral
representation of the cut quark propagator (see, e.g., Sec.~6.3 of
\cite{D'Hoker:2004aa} and Sec.~2.7.2 of \cite{Romao:2013aa}),
\begin{align} 
  \Xi(l) =  
  \int d \mu^2 \big[ J_1 (\mu^2)\,\mu\,\id + J_2 (\mu^2)\,\lslash \big] \,
  \delta(l^2 -\mu^2) \ ,
\label{e:jetspectral}
\end{align}
where $\mu^2$ can be interpreted as the invariant mass of the current jet, {\it
  i.e.}, of the particles going through the cut in the top blob of
Fig.\ref{fig:handbags}(a). The $J_i$ are the spectral functions of the
quark propagator, also called ``jet functions'' in
\cite{Accardi:2008ne}, and can be interpreted as current-jet mass distributions. As a consequence of positivity constraints and CPT invariance, These satisfy \cite{D'Hoker:2004aa,Romao:2013aa,Weinberg:1995mt}
\begin{align}
  J_2(\mu^2) \geq J_1(\mu^2) \geq 0
  \hspace*{0.5cm} \text{and} \hspace*{0.5cm}
  \int d\mu^2 J_2(\mu^2) = 1 \ .
\label{eq:jetfnsprops}
\end{align}
From a comparison of Eqs.\eqref{e:jetexpansion} and \eqref{e:jetspectral}, one can see that 
\begin{align}
  A_1(l^2)&=\frac{\sqrt{l^2}}{\Lambda}J_1(l^2) & A_2(l^2)&=J_2(l^2) \ .
  \label{eq:jet_vs_spectral}
\end{align}

\begin{figure*}[bt]
  \centering
  (a)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram0}
  \hfill
  (b)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram2}
  \hfill
  (c)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram1}
  \caption{Diagrams contributing to inclusive DIS scattering up to twist-3, including a jet correlator in the top part. Note the gluon attaches to both the nucleon and jet correlators. The Hermitian conjugates of diagrams (b) and (c), i.e., with gluons attaching to the right of the cut, are not shown.
  }
  \label{fig:handbags}
\end{figure*}

@@

Finally, the contraction of the hadronic tensor with the leptonic tensor leads
to the following result for the inclusive DIS cross section up to order $\Lambda/Q$~\cite{Bacchetta:2006tn}:
\begin{align}
\frac{d\sigma}{d\xbj \, dy\, d\phi_S}
%&
=
\frac{2 \alpha^2}{\xbj y Q^2}\,
\frac{y^2}{2\,(1-\varepsilon)}\, 
\biggl\{
&F_{T} + \varepsilon F_{L}
+ S_\parallel \lambda_e\,
  \sqrt{1-\varepsilon^2}\; 
F_{LL}
%\nonumber 
%\\  &
+ |\bm{S}_\perp| \lambda_e\, \sqrt{2\,\varepsilon (1-\varepsilon)}\, 
  \cos\phi_S\, 
F_{LT}^{\cos \phi_S}
 \biggr\} \ ,
\label{e:crossdis}
\end{align}
where $\phi_S$ is the angle between the transverse component of the proton spin vector and the lepton plane, $\epsilon$ is the ratio of the longitudinal and transverse photon fluxes, and $\lambda_e$ is the electron's helicity. 
The structure functions on the right hand side read
\begin{align}
F_{UU ,T} &= \xbj\,\sum_q e_q^2\,f_1^q(\xbj),
\\
F_{UU ,L} &= 0,
\\
F_{LL} &=\xbj\,\sum_q e_q^2\,g_1^q(\xbj),
\label{e:FLLint}
\\
F_{UT}^{\sin \phi_S}&=0,
\label{e:FUTint}
\\
F_{LT}^{\cos \phi_S}&=-\xbj\,\sum_q e_q^2\, \frac{2\Lambda}{Q}\,
\biggl(\xbj  g_T^q(\xbj)
   + \frac{\mj -\mq}{\Lambda} \, h_{1}^q(\xbj) \biggr) \ ,
\label{e:FLTint}
\end{align}
where $f_1^q$, $g_1^q$ and $h_1^q$ are the unpolarized, polarized, and transversity PDFs,respectively.
The second term in the last structure function is a new result from our
analysis, it is proportional to the jet mass, and it is not suppressed as an
inverse power of $Q$ compared to the 
standard $g_T$ term. Perturbatively, $M_q^\text{pert}=m_q$ and the new term vanishes. However, on the nonperturbative vacuum the jet mass $M_q$ is much larger than the quark's current mass $m_q$, originating a nonnegligible term to the twist-3 part of the target's $g_2$ structure function, as we will discuss in the next section.

@@

The structure functions in Eqs.~\eqref{e:FLLint} -- \eqref{e:FLTint} can be
related to the usual
structure functions $g_1$ and $g_2$ defined. In
particular, the expression of the structure function $g_2$ is
\begin{equation} 
g_2=\frac12 \sum_q e_q^2(g_T^q-g_1^q)
\end{equation}
structure function. Following the derivation in Ref.~\cite{Accardi:2009au}, one finds
\begin{align}
\label{e:g2}
  g_2(\xbj) = g_2^{WW}(\xbj) + \frac{1}{2}\,\sum_a e_a^2
\biggl(
    \widetilde g_T^{a \star}(\xbj) 
    + \int_\xbj^1\frac{dy}{y} \widehat{g}_T^q(y) 
    + \frac{\mq}{\Lambda} \left(\frac{h_1^q}{x}\right)^\star(\xbj) 
    + \frac{\mj-\mq}{\Lambda} \frac{h_1^q(\xbj)}{\xbj} 
\Biggr) \ ,
\end{align}
where we defined $f^*(x) = -f(x) + \int_x^1\frac{dy}{y} f(y)$, and $\tilde g_T$ and $\hat g_T$ are pure twist-3 functions defined in that reference. The first four terms coincide with the result obtained in the conventional handbag
approximation \cite{Accardi:2009au}, while the last is new. Note that even if
the relation is written for the sum over quark flavors weighted by their charge
squared, it is also valid flavor by flavor; in fact, the steps
leading to such a decomposition are formulated at the quark correlator level.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\includegraphics[width=8cm]{g2contribN}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left)
and neutron (right) $g_2$ structure functions compared to the JAM15 fit of the
$g_2-g_2^{\text{WW}}$ function (solid black) \cite{Sato:2016tuz}. The quark and jet
contributions are shown with a dotted red and a dot-dashed green line
respectively, with uncertainty bands coming form the Pavia15 fit of the
transversity function \cite{Radici:2015mwa}. The uncertainty in the choice
$m_q=5$ MeV and $\mj=100$ MeV is not shown. The pure twist-3 contribution
calculated by Braun et al. \cite{Braun:2011aw} is shown as a dashed blue line
(no uncertainty estimate was provided in the original reference). 
}
\end{center}
\end{figure}

@@

It is interesting to consider the moments of the non Wandzura-Wilczek contribution to $g_2$,
\begin{align}
  d_N \equiv (N+1) \int_0^1 x^N \bigg( g_2(x) - g_2^{WW}(x) \bigg) \ .
\end{align}
For a generic function $f$, let us define it's $N$-th moment as $f[N]=\int_0^1 dx\, x^{N} f(x)$. It is then straightforward to verify that $f^*[N] = - f[N]\,N/(N+1)$ and  
\begin{align}
  d_N & = (N+1) g_2[N] + N g_1[N] \\
  & = \frac12 \sum_q e_q^2 \bigg[ \hat g_T^q[N] - N \tilde g_T^q[N] 
    + \frac{(N+1) \mj-(2N+1)\mq}{\Lambda} h_1^q[N-1] \bigg] \ .
\end{align}

The zero-th moment, $d_0=\int g_2$, provides an interesting relationship between transversity and the inclusive structure function $g_2$:
\begin{align}
  \label{eq:BC}
  \int dx\, g_2(x) = \sum_q e_q^2 \frac{\mj-\mq}{\Lambda} \int dx\, \frac{1}{x} h_1^q(x) \ .
\end{align}
Here we used the fact that $\widehat g_T^q[0]$ vanishes identically due to the symmetry properties of the quark-gluon-quark correlators \cite{Accardi:2009au}.
%, or equivalently as a consequence of the Lorentz invariance of QCD interactions, that entails $\int_0^1 dx g_1^a(x) = \int_0^1 g_T^q(x)$. 
%
Note that the only surviving term on the right-hand side is the new jet contribution. Thus, our new sum rule \eqref{eq:BC}
generalizes the Burkhardt--Cottingham (BC) sum rule~\cite{Burkhardt:1970ti},
which states that  $\int_0^1 dx\, g_2(x) =0$, while we have shown that jet-mass corrections, and in particular from invariant mass generation in spin-flip processes,  can directly violate this.
In fact, the possibility of a violation of the BC sum rule due to 
contributions from spin-flip processes was already mentioned in the original
derivation, but these do not show up in treatments that only
consider free-field quark propagators for the struck quark
\cite{Jaffe:1996zw}.  Although we formulated \eqref{eq:BC} in terms of a sum over quark flavors in order to display a clear
connection to the structure function $g_2$, we stress that this is valid also
flavor by flavor, i.e., for each single flavor the only measurable nonzero contribution to the zeroth moment of the
structure function $g_2^q$ comes from the coupling between its jet mass and transversity function\footnote{This conclusion is true even if the BC sum rule is broken by a $J = 0$ fixed pole with
non-polynomial residue \cite{Jaffe:1996zw}, since this would appear as a
$\delta(\xbj)$ contribution and would not be measurable.}.

One should notice that since $h_1$ is slowly driven to 0 by QCD evolution as $Q^2 \to \infty$, the BC sum rule is still satisfied at least
asymptotically.
At finite scales, however, the only way to preserve the validity of the
BC sum rule is if
%that is derived at the cross section level (under suitable assumptions on the
%high-energy behavior of te hspin-flip amplitude) rather than from an analysis
%of collinear correlators, as done here, 
%to be strictly valid, we obtain a
%strong constraint on the transversity function, 
\begin{align}
   \int dx\, \frac{1}{x} h_1^q(x) = 0 \ .
\label{eq:ABsumrule}
\end{align}
Interestingly, one can show that this constraint, if valid at any given scale
$Q_0$ is conserved through QCD evolution. However, we think that it is unlikely to be satisfied in general, since Eq.\eqref{eq:ABsumrule} is explicitly broken in perturbative QCD~\cite{Kundu:2001pk}, as well as in model calculations \cite{Schweitzer:2001sr,Wakamatsu:2007nc,Pasquini:2005dk,Cloet:2007em,Bacchetta:2008af,Bourrely:2010ng}. Then, if we assume that the BC sum rule is broken by a {\em finite} amount (which is very reasonable from the recent small-$x$ estimate $g_2\propto x^{\sqrt{ \alpha_s N_c / \pi}}$, with $N_c=3$ the number of colors, and $\alpha_s$ the strong coupling constant \cite{Kovchegov:2016zex})
we obtain that $h_1(x)/x$ must be integrable. This implies a 
a bound on the small $x$ behavior of the transversity, 
\begin{align}
  h_1^q(x) \propto x^\epsilon \ \ \ \epsilon>0 \ .
\label{eq:ABbound}
\end{align}
This bound can be very useful, \eg, in transversity fits, where the data at
small $x$ is, as yet, very limited, and in general for proper extrapolations
to $x_B=0$ when experimentally measuring moments.

The first moment, $d_1$, is the first one to display a contribution from the pure twist-3 part of $g_2$:
\begin{align}
  d_1 & = \frac12 \sum_q e_q^2 \bigg(  \hat g_T^q[1] - 2 \tilde g_T^q[1]
    + \frac{2\mj-3\mq}{\Lambda} h_1^q[0] \bigg)
\label{e:d1}
\end{align}
where $h_1^q[0] = \int_0^1 dx h_1^q(x)$ is the contribution of a quark $q$ to the target's tensor charge. The second moment,
\begin{align}
  d_2 & = \frac12 \sum_q e_q^2 \bigg(\hat g_T^q[2] - 3 \tilde g_T^q[2] 
    + \frac{3\mj-4\mq}{\Lambda} h_1^q[1] \bigg) \ ,
\end{align}
is also interesting because the pure twist-3 part can be related to quark-gluon-quark local matrix elements, see \cite{Jaffe:1996zw}, and
interpreted as the average color force experienced by the struck quark as
it exits the nucleon \cite{Burkardt:2012sd}; for experimental
measurements of $d_2$, see, e.g., Refs.~\cite{Anthony:2002hy,Slifer:2008xu,Solvignon:2013yun,Posik:2014usi,Flay:2016wie}.

@@

\begin{figure}[tbh]
  \centering
  \parbox[c]{8cm}{\vskip.1cm\centering
  \includegraphics
      [width=7cm,clip=true]
      {epl_emn_h_X}
  }
  \parbox[c]{8cm}{\vskip-0.2cm\centering
  \includegraphics
      [width=7cm,clip=true]
      {epl_emn_h_h_X}
  }
  \caption{\small
	Single hadron {\it (left)} and double hadron {\it (right)}
	production in $e^+ e^-$ collisions at LO with jet and
	fragmentation correlators.
  }
  \label{fig:epl_emn_h}
\end{figure}

To conclude this section, we note that the jet contribution also leads to an explicit breaking of the Efremov--Teryaev--Leader (ETL) sum rule \cite{Efremov:1996hd}, in which the pure twist-3 contribution to the first moment of $g_2-g_2^{WW}$ also disappears. To see this, let's define the valence contribution to a given structure function as $f^V=\frac12\sum_q e_q^2 (f^q-f^{\bar q})$. Then, as shown in \cite{Efremov:1996hd}, $\widehat g_T^V[1] - 2\tilde g_T^V[1] = 0$, and from Eq.~\eqref{e:d1} we obtain
\begin{align}
  d_1^V = \frac12 \sum_q e_q^2 \frac{2\mj- \mq}{\Lambda} 
    \big(h_1^{q}[0]-h_1^{\bar q}[0]\big) \ .
\end{align}
Assuming $M_\text{light}\equiv M_u \approx M_d \gg m_u , m_d$ and isospin symmetry of the proton and neutron, we can also see that 
\begin{align}
  d_1^V = \frac{M_{\text{light}}}{\Lambda} \delta_T(p) \ ,   
\end{align}
This gives an alternative way to access the proton tensor charge, $\delta_T(p) = \sum_q e_q^2 \big(h_1^{q}[0]-h_1^{\bar q}[0]\big)$, by measuring or fitting moments of the flavor separated $g_2$ structure function.
%\AAcom{I tried to build an observable for $d_1^V$ such as $p-n$ or $p-\bar p$, but I am messing up a bit.}

>>>>

\preprint{JLAB-THY-17-2483}

@@

In this paper, we discuss a novel way of accessing the transversity parton distribution function (PDF) and measuring the proton's tensor charge in totally inclusive Deep Inelastic Scattering. 
We revisit the standard analysis of the DIS handbag diagram, taking into account the fact that on-shell quarks cannot, in fact, be present in the final state, but they rather decay and form (mini)jets of hadrons. This is sufficient to modify the structure of the DIS cut diagram, even if none of those hadrons is detected in the final state. For a proper description of this effect, we include ``jet correlators'' into the analysis, and pay particular attention to ensuring that our results are gauge invariant.

@@

The correlator $\Xi$ can be parametrized in terms of jet parton correlation
functions $A_i$ and $B_i$ through a Lorentz covariant Dirac decomposition that utilizes the vectors $l$ and $n_+$, 
\begin{equation}
\Xi(l,n_+) = \Lambda A_1(l^2)\,\id + A_2(l^2)\,\lslash 
+ \frac{\Lambda^2}{l \cdot n_{+}} \nslash_+ \, B_{1}(l^2)
+ \frac{i \Lambda}{2 l \cdot n_{+}} [\,\lslash,\nslash_+ ] \, B_{2}(l^2) \ ,
\label{e:jetexpansion}
\end{equation} 
where $\Lambda$ is an arbitrary scale, introduced for power counting purposes.
In this parametrization, no terms proportional to $\gamma_5$ enter because of parity invariance. Time reversal invariance in QCD requires $B_{2}=0$, while $B_{1}$ contributes only at twist-4 order and will not be considered further in this paper. We focus, instead, on the role of chiral odd terms in the $g_2$ structure function up to twist 3. At this order, 
\begin{equation}
  \Xi(l,n_+) = \Lambda A_1(l^2)\,\id + A_2(l^2)\,\lslash 
    + {\cal O}(\Lambda^2/Q^2)
\label{e:jetexpansion-tw3}
\end{equation} 
is nothing else than the cut quark propagator; note however, that we consider
here the full QCD vacuum rather than the perturbative one (or, in other words, the interacting rather than the free quark fields).
The $A_1$ and $A_2$ terms can be interpreted in terms of the spectral
representation of the cut quark propagator (see, e.g., Sec.~6.3 of
\cite{D'Hoker:2004aa} and Sec.~2.7.2 of \cite{Romao:2013aa}),
\begin{align} 
  \Xi(l) =  
  \int d \mu^2 \big[ J_1 (\mu^2)\,\mu\,\id + J_2 (\mu^2)\,\lslash \big] \,
  \delta(l^2 -\mu^2) \ ,
\label{e:jetspectral}
\end{align}
where $\mu^2$ can be interpreted as the invariant mass of the current jet, {\it
  i.e.}, of the particles going through the cut in the top blob of
Fig.\ref{fig:handbags}(a). The $J_i$ are the spectral functions of the
quark propagator, also called ``jet functions'' in
\cite{Accardi:2008ne}, and can be interpreted as current-jet mass distributions. As a consequence of positivity constraints and CPT invariance, these satisfy \cite{D'Hoker:2004aa,Romao:2013aa,Weinberg:1995mt}
\begin{align}
  J_2(\mu^2) \geq J_1(\mu^2) \geq 0
  \hspace*{0.5cm} \text{and} \hspace*{0.5cm}
  \int d\mu^2 J_2(\mu^2) = 1 \ .
\label{eq:jetfnsprops}
\end{align}
From a comparison of Eqs.\eqref{e:jetexpansion} and \eqref{e:jetspectral}, one can see that 
\begin{align}
  A_1(l^2)&=\frac{\sqrt{l^2}}{\Lambda}J_1(l^2) & A_2(l^2)&=J_2(l^2) \ .
  \label{eq:jet_vs_spectral}
\end{align}

\begin{figure*}[bt]
  \centering
  (a)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram0.eps}
  \hfill
  (b)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram2.eps}
  \hfill
  (c)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram1.eps}
  \caption{Diagrams contributing to inclusive DIS scattering up to twist-3, including a jet correlator in the top part. Note the gluon attaches to both the nucleon and jet correlators. The Hermitian conjugates of diagrams (b) and (c), i.e., with gluons attaching to the right of the cut, are not shown.
  }
  \label{fig:handbags}
\end{figure*}

@@

Finally, the contraction of the hadronic tensor with the leptonic tensor leads
to the following result for the inclusive DIS cross section up to order $\Lambda/Q$~\cite{Bacchetta:2006tn}:
\begin{align}
\frac{d\sigma}{d\xbj \, dy\, d\phi_S}
%&
=
\frac{2 \alpha^2}{\xbj y Q^2}\,
\frac{y^2}{2\,(1-\varepsilon)}\, 
\biggl\{
&F_{T} + \varepsilon F_{L}
+ S_\parallel \lambda_e\,
  \sqrt{1-\varepsilon^2}\; 
F_{LL}
%\nonumber 
%\\  &
+ |\bm{S}_\perp| \lambda_e\, \sqrt{2\,\varepsilon (1-\varepsilon)}\, 
  \cos\phi_S\, 
F_{LT}^{\cos \phi_S}
 \biggr\} \ ,
\label{e:crossdis}
\end{align}
where $\phi_S$ is the angle between the transverse component of the proton spin vector and the lepton plane, $\epsilon$ is the ratio of the longitudinal and transverse photon fluxes, and $\lambda_e$ is the electron's helicity. 
The structure functions on the right hand side read
\begin{align}
F_{T} &= \xbj\,\sum_q e_q^2\,f_1^q(\xbj),
\\
F_{L} &= 0,
\\
F_{LL} &=\xbj\,\sum_q e_q^2\,g_1^q(\xbj),
\label{e:FLLint}
\\
F_{UT}^{\sin \phi_S}&=0,
\label{e:FUTint}
\\
F_{LT}^{\cos \phi_S}&=-\xbj\,\sum_q e_q^2\, \frac{2\Lambda}{Q}\,
\biggl(\xbj  g_T^q(\xbj)
   + \frac{\mj -\mq}{\Lambda} \, h_{1}^q(\xbj) \biggr) \ ,
\label{e:FLTint}
\end{align}
where $f_1^q$, $g_1^q$ and $h_1^q$ are the unpolarized, polarized, and transversity PDFs,respectively.
The second term in the last structure function is a new result from our
analysis, it is proportional to the jet mass, and it is not suppressed as an
inverse power of $Q$ compared to the 
standard $g_T$ term. Perturbatively, $M_q^\text{pert}=m_q$ and the new term vanishes. However, on the nonperturbative vacuum the jet mass $M_q$ is much larger than the quark's current mass $m_q$, originating a nonnegligible term to the twist-3 part of the target's $g_2$ structure function, as we will discuss in the next section.

@@

The structure functions in Eqs.~\eqref{e:FLLint} -- \eqref{e:FLTint} can be
related to the usual
structure functions $g_1$ and $g_2$ defined from the following Lorentz decomposition of the hadronic tensor
\begin{align}
\begin{split}
W^{\mu\nu}(P,q) 
  & = \frac{1}{P\cdot q} \varepsilon^{\mu\nu\rho\sigma} q_\rho 
  \Big[ S_\sigma g_1(x_B,Q^2) 
    + \Big( S_\sigma - \frac{S\cdot q}{P\cdot q}\, p_\sigma
	\Big) g_2(x_B,Q^2)
  \Big] \ ,
\label{eq:Wmunu}
\end{split}
\end{align}
Then, in particular,
\begin{align}
  g_1 &= \frac{1}{x_B} F_{LL} \\
  g_2+g_1 &= \frac{Q}{2\Lambda x_B^2} F_{LT}^{\cos \phi_S} \ .
\end{align}
% \begin{align}
% g_1&=\frac12 \sum_q e_q^2g_1^q \ , \\
% g_2&=\frac12 \sum_q e_q^2(g_T^q-g_1^q) \ .
% \end{align}
Utilizing Equations of Motion and Lorentz Invariance relations as discussed in Ref.~\cite{Accardi:2009au}, we arrive at
\begin{align}
\label{e:g2}
  g_2(\xbj) = g_2^{WW}(\xbj) + \frac{1}{2}\,\sum_a e_a^2
\biggl(
    \widetilde g_T^{a \star}(\xbj) 
    + \int_\xbj^1\frac{dy}{y} \widehat{g}_T^q(y) 
    + \frac{\mq}{\Lambda} \left(\frac{h_1^q}{x}\right)^\star(\xbj) 
    + \frac{\mj-\mq}{\Lambda} \frac{h_1^q(\xbj)}{\xbj} 
\Biggr) \ ,
\end{align}
where we used $f^*(x) = f(x) - \int_x^1\frac{dy}{y} f(y)$, and $\tilde g_T$ and $\hat g_T$ are pure twist-3 functions that only depend on projections of quark-gluon-quark correlator, and are explicitly defined in that reference. The first four terms coincide with the result obtained in the conventional handbag
approximation \cite{Accardi:2009au}, while the last is new. Note that even if
the relation is written for the sum over quark flavors weighted by their charge
squared, it is also valid flavor by flavor; in fact, the steps
leading to such a decomposition are formulated at the quark correlator level.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8cm]{g2contrib.pdf}
\includegraphics[width=8cm]{g2contribN.pdf}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left)
and neutron (right) $g_2$ structure functions compared to the JAM15 fit of the
$g_2-g_2^{\text{WW}}$ function (solid black) \cite{Sato:2016tuz}. The quark and jet
contributions are shown with a dotted red and a dot-dashed green line
respectively, with uncertainty bands coming form the Pavia15 fit of the
transversity function \cite{Radici:2015mwa}. The uncertainty in the choice
$m_q=5$ MeV and $\mj=100$ MeV is not shown. The pure twist-3 contribution
calculated by Braun et al. \cite{Braun:2011aw} is shown as a dashed blue line
(no uncertainty estimate was provided in the original reference). 
}
\end{center}
\end{figure}

@@

It is interesting to consider the moments of the non Wandzura-Wilczek contribution to $g_2$,
\begin{align}
  d_N \equiv (N+1) \int_0^1 dx\,x^N \bigg( g_2(x) - g_2^{WW}(x) \bigg) \ .
\end{align}
For a generic function $f$, let us define it's $N$-th moment as $f[N]=\int_0^1 dx\, x^{N} f(x)$. It is then straightforward to verify that $f^*[N] = f[N]\,N/(N+1)$ and  
\begin{align}
  d_N & = (N+1) g_2[N] + N g_1[N] \\
  & = \frac12 \sum_q e_q^2 \bigg[ N \tilde g_T^q[N] + \hat g_T^q[N]  
    + \frac{(N+1) \mj- \mq}{\Lambda} h_1^q[N-1] \bigg] \ .
\end{align}

The zero-th moment, $d_0=\int g_2$, provides an interesting relationship between transversity and the inclusive structure function $g_2$:
\begin{align}
  \label{eq:BC}
  \int dx\, g_2(x) = \sum_q e_q^2 \frac{\mj-\mq}{\Lambda} \int dx\, \frac{1}{x} h_1^q(x) \ .
\end{align}
Here we used the fact that $\widehat g_T^q[0]$ vanishes identically due to the symmetry properties of the quark-gluon-quark correlators \cite{Accardi:2009au}.
%, or equivalently as a consequence of the Lorentz invariance of QCD interactions, that entails $\int_0^1 dx g_1^a(x) = \int_0^1 g_T^q(x)$. 
%
Note that all pure twist-3 terms have explicitly disappeared, and the only surviving term on the right-hand side is the new jet contribution. Thus, our new sum rule \eqref{eq:BC}
generalizes the Burkhardt--Cottingham (BC) sum rule~\cite{Burkhardt:1970ti},
which states that  $\int_0^1 dx\, g_2(x) =0$, while we have shown that jet-mass corrections, and in particular from invariant mass generation in spin-flip processes, can directly violate this.
In fact, the possibility of a violation of the BC sum rule due to 
contributions from spin-flip processes was already mentioned in the original
derivation, but these do not show up in treatments that only
consider free-field quark propagators for the struck quark
\cite{Jaffe:1996zw}.  Although we formulated \eqref{eq:BC} in terms of a sum over quark flavors in order to display a clear
connection to the structure function $g_2$, we stress that this is valid also
flavor by flavor, i.e., for each single flavor the only measurable nonzero contribution to the zeroth moment of the
structure function $g_2^q$ comes from the coupling between its jet mass and transversity function\footnote{This conclusion is true even if the BC sum rule is broken by a $J = 0$ fixed pole with
non-polynomial residue \cite{Jaffe:1996zw}, since this would appear as a
$\delta(\xbj)$ contribution and would not be measurable.}.

One should notice that since $h_1$ is slowly driven to 0 by QCD evolution as $Q^2 \to \infty$, the BC sum rule is still satisfied at least
asymptotically.
At finite scales, however, the only way to preserve the validity of the
BC sum rule is if
%that is derived at the cross section level (under suitable assumptions on the
%high-energy behavior of te hspin-flip amplitude) rather than from an analysis
%of collinear correlators, as done here, 
%to be strictly valid, we obtain a
%strong constraint on the transversity function, 
$
%\begin{align}
   \int dx\, \frac{1}{x} h_1^q(x) = 0 \ .
\label{eq:ABsumrule}
%\end{align}
$
Interestingly, one can show that this constraint, if valid at any given scale
$Q_0$ is conserved through QCD evolution. However, we think that it is unlikely to be satisfied in general, since it is explicitly broken in perturbative QCD~\cite{Kundu:2001pk}, as well as in model calculations \cite{Schweitzer:2001sr,Wakamatsu:2007nc,Pasquini:2005dk,Cloet:2007em,Bacchetta:2008af,Bourrely:2010ng}. Then, if we assume that the BC sum rule is broken by a {\em finite} amount we obtain that $h_1(x)/x$ must be integrable, which implies the following bound on the small $x$ behavior of the transversity: 
\begin{align}
  h_1^q(x) \propto x^\epsilon \ \ \ \epsilon>0 \ .
\label{eq:ABbound}
\end{align}
This bound can be very useful, \eg, in transversity fits, where the data at
small $x$ is, as yet, very limited, and in general for proper extrapolations
to $x_B=0$ when experimentally measuring moments. We also note that the small-$x$ behavior of the longitudinal spin structure function $g_1$ has been recently investigated in Ref.~\cite{Kovchegov:2016zex}; it would be interesting to extend those techniques to the transverse spin structure functions $g_T=g_1+g_2$ and $h_1$.

The first moment, $d_1$, is the first one to display a contribution from the pure twist-3 part of $g_2$:
\begin{align}
  d_1 & = \frac12 \sum_q e_q^2 \bigg( 2 \tilde g_T^q[1] + \hat g_T^q[1] 
    + \frac{2\mj-\mq}{\Lambda} h_1^q[0] \bigg)
\label{e:d1}
\end{align}
where $h_1^q[0] = \int_0^1 dx h_1^q(x)$ is the contribution of a quark $q$ to the target's tensor charge. The second moment,
\begin{align}
  d_2 & = \frac12 \sum_q e_q^2 \bigg( 3 \tilde g_T^q[2] + \hat g_T^q[2] 
    + \frac{3\mj-\mq}{\Lambda} h_1^q[1] \bigg) \ ,
\end{align}
is also interesting because the pure twist-3 part can be related to quark-gluon-quark local matrix elements, see \cite{Jaffe:1996zw}, and
interpreted as the average color force experienced by the struck quark as
it exits the nucleon \cite{Burkardt:2012sd}; for experimental
measurements of $d_2$, see, e.g., Refs.~\cite{Anthony:2002hy,Slifer:2008xu,Solvignon:2013yun,Posik:2014usi,Flay:2016wie}.

@@

\begin{figure}[tbh]
  \centering
  \parbox[c]{8cm}{\vskip.1cm\centering
  \includegraphics
      [width=7cm,clip=true]
      {epl_emn_h_X.eps}
  }
  \parbox[c]{8cm}{\vskip-0.2cm\centering
  \includegraphics
      [width=7cm,clip=true]
      {epl_emn_h_h_X.eps}
  }
  \caption{\small
	Single hadron {\it (left)} and double hadron {\it (right)}
	production in $e^+ e^-$ collisions at LO with jet and
	fragmentation correlators.
  }
  \label{fig:epl_emn_h.eps}
\end{figure}

To conclude this section, we note that the jet contribution also leads to an explicit breaking of the Efremov--Teryaev--Leader (ETL) sum rule \cite{Efremov:1996hd}, in which the pure twist-3 contribution to the first moment of $g_2-g_2^{WW}$ also disappears. To see this, let's define the valence contribution to a given structure function as $f^V=\frac12\sum_q e_q^2 (f^q-f^{\bar q})$. Then, as shown in \cite{Efremov:1996hd}, $\widehat 2\tilde g_T^V[1] + g_T^V[1]  = 0$, and from Eq.~\eqref{e:d1} we obtain
\begin{align}
  d_1^V = \frac12 \sum_q e_q^2 \frac{2\mj- \mq}{\Lambda} 
    \big(h_1^{q}[0]-h_1^{\bar q}[0]\big) \ .
\end{align}
Assuming $M_\text{light}\equiv M_u \approx M_d \gg m_u , m_d$ and isospin symmetry of the proton and neutron, we can also see that 
\begin{align}
  d_1^V = \frac{M_{\text{light}}}{\Lambda} \delta_T(p) \ ,   
\end{align}
This gives an alternative way to access the proton tensor charge, $\delta_T(p) = \sum_q e_q^2 \big(h_1^{q}[0]-h_1^{\bar q}[0]\big)$, by measuring or fitting moments of the flavor separated $g_2$ structure function.
%\AAcom{I tried to build an observable for $d_1^V$ such as $p-n$ or $p-\bar p$, but I am messing up a bit.}
